---
layout: post
title:  "tm.plugin.webmining 1.3 on CRAN"
categories: [Textmining, tm.plugin.webmining]
tags: [R, Release]
---

**tm.plugin.webmining** supports the retrieval of full-text content from 
structured news feeds (e.g. RSS) and HTML pages in a 2-step procedure:

1. Download meta data from the feed (through *WebSource*)
2. Download and extract main content for the feed item (through *WebCorpus*)

The newest version of **tm.plugin.webmining** 1.3 has now been 
released on [CRAN](http://cran.r-project.org/web/packages/tm.plugin.webmining/index.html) including minor changes and bug-fixes:

- Fix [issue](https://github.com/mannau/tm.plugin.webmining/issues/7) with Yahoo 
  News Source; since Yahoo News seems to have de-activated RSS feeds we now switched to 
  HTML retrieval. Thanks Eliano for reporting!
- Fix retrieval [issue](https://github.com/mannau/tm.plugin.webmining/issues/6) 
  with NYTimes Source.

<!--more-->

To retrieve current financial news for Tesla (TSLA) from Google Finance we can 
use the following code:

```{r web-1, results='hide', message=FALSE, warning=FALSE}
library(tm)
library(tm.plugin.webmining)
```

The resulting corpus of length 20 includes the following meta-data:
```{r web-2, cache = TRUE, warning = FALSE}
gf <- WebCorpus(GoogleFinanceSource("NASDAQ:TSLA"))
meta(gf[[1]])
```

We can now show a frequency histogram of datetimestamps as follows:
```{r web-3}
gf <- WebCorpus(GoogleFinanceSource("NASDAQ:TSLA"))
hist(do.call(c, lapply(gf, meta, "datetimestamp")), breaks = "days", 
freq = TRUE, col = "grey")
```

The corpus is not big enough for serious statistical analysis, but should be 
sufficient for a simple wordcloud:
```{r web-4, results='hide', message=FALSE, warning=FALSE}
library(wordcloud)
```

Remove stopwords and show wordcloud:
```{r web-5}
gf.proc <- tm_map(gf, content_transformer(tolower))
gf.proc <- tm_map(gf.proc, removeWords, stopwords("english"))
wordcloud(gf.proc, colors=brewer.pal(6,"Dark2"),random.order=FALSE)
```
