---
layout: post
title:  Working with HDF5 files in R using h5
categories: [HDF5, h5]
tags: [R, Release]
---

After quite some work I'm happy to announce the CRAN release of 
**[h5](http://cran.r-project.org/web/packages/h5/index.html)** - an R 
interface to the [HDF5](https://www.hdfgroup.org/HDF5/) library. It utilizes the 
[HDF5 C++ API](https://www.hdfgroup.org/HDF5/doc/cpplus_RM/) through 
**[Rcpp](http://cran.r-project.org/web/packages/Rcpp/index.html)** and S4 classes. 
Although the package is still under heavy development with some 
[issues](https://github.com/mannau/h5/issues) remaining it is already covered by 
200+ test cases and runs on all major platforms (Windows, OS X, Linux). I 
started working on the package after experiencing some speed and usability issues 
with existing packages (e.g. **[rhdf5](http://www.bioconductor.org/packages/release/bioc/html/rhdf5.html)**). 

<!--more-->

[HDF5](https://www.hdfgroup.org/HDF5/) is an excellent library and data model to 
store huge amounts of data in a binary file format. Supporting most major 
platforms and programming languages it can be used to exchange data files in a 
language independent format. Compared to R's integrated *save()* and *load()* 
functions it also supports access to only parts of the binary data files and can
therefore be used to process data not fitting into memory.

Now it's time for a small **h5** use case. First we create an HDF5 file and store
a numeric vector, an integer matrix and a character array.

```{r h5-0, results='hide', echo=FALSE}
library(methods)
if(file.exists("test.h5")) {
	file.remove("test.h5")
}
```

```{r h5-1, results='hide', message=FALSE, warning=FALSE}
library(h5)
testvec <- rnorm(10)
testmat <- matrix(1:9, nrow = 3)
row.names(testmat) <- 1:3
colnames(testmat) <- c("A", "BE", "BU")
letters1 <- paste(LETTERS[runif(45, min = 1, max = length(LETTERS))])
letters2 <- paste(LETTERS[runif(45, min = 1, max = length(LETTERS))])
testarray <- array(paste0(letters1, letters2), c(3, 3, 5))

file <- h5file("test.h5")
# Save testvec in group 'test' as DataSet 'testvec'
file["test", "testvec"] <- testvec
file["test", "testmat"] <- testmat
file["test", "testarray"] <- testarray
h5close(file)
```

We can now retrieve the data from the file

```{r h5-2, results='show', message=FALSE, warning=FALSE}
file <- H5File("test.h5")
dataset_testmat <- file["test", "testmat"]
# We can now retrieve all data from the DataSet object using e.g. the  subsetting operator
dataset_testmat[]

# We can also subset the data directly, e.g. row 1 and 3
dataset_testmat[c(1, 3), ]
```

Note, that we have now lost the row- and column names associated with the *testmat* object
in the retrieved matrix. HDF5 supports metadata with attributes, which we need to
add to (retrieve from) the DataSet manually.

```{r h5-3, results='hide', message=FALSE, warning=FALSE}
h5attr(dataset_testmat, "rownames") <- row.names(testmat)
h5attr(dataset_testmat, "colnames") <- colnames(testmat)
```

We can now retrieve our matrix including meta-data as follows:

```{r h5-4, results='show', message=FALSE, warning=FALSE}
outmat <- dataset_testmat[]
row.names(outmat) <- h5attr(dataset_testmat, "rownames")
colnames(outmat) <- h5attr(dataset_testmat, "colnames")
identical(outmat, testmat)
```

Do not forget to close the HDF5 file in the end

```{r h5-5, results='hide', message=FALSE, warning=FALSE}
h5close(file)
```

You can find many more code examples as unit tests in the **h5** package.

The next blog post will cover selection and hyperslabs in more detail, stay
tuned... 
